Delivering Reliable Software in the European Scientific Arena: Evolution of Methodologies and New Trend Analysis

### 2018-01-09 paper-ieee-trans.tex

The DataGrid \cite{cordis:datagrid} project (Jan 2001 -- Dec 2003) 
had as main goal to develop the software to provide basic grid functionality
and associated management tools for a large scale testbed for demonstration projects in three
specific areas of science, Particle physics, Bioinformatics and Earth observation
%was devoted to develop the first European infrastructure for Grid computing. 
The project brought together 21 academic and industry partners, from 15 different
countries. 
%There were 3 application areas for requirement gathering; Particle physics, Bioinformatics and Earth observation. 
The project delivered its own software distribution, named EDG (EU DataGrid), strongly 
based on Globus Middleware Services \cite{globus}. The project was organized in 12 workpages, from 
which 5 were devoted to software development coordinated by an 
Architecture Task Force, supervising the overall design and technical consistency 
of the developments.

@@

The source code was available at a private, centralized version control system.
The code was not validated by any Quality Assurance policy, but passed through a manual
certification procedure at the time of the release. This procedure tried to
improve the reliability of the software components by applying a set of
requirements [ref] to be fulfilled at each stage of the  development lifecycle (integration, certification,
pre--production, production), carried out by different teams. However, with 
EGEE--II, the developers started to use the the ETICS system \cite{etics}, solution aiming to develop
an automatic build system for Grid middleware.

@@

EGI--InSPIRE not only ensured the coordination of the operation of a secure, reliable 
European-wide production grid infrastructure federated from national grid initiatives 
that is integrated and interoperates with other grids worldwide but also foreseed the 
maintenance and development of operational tools, like the Operations Portal \cite{egi-ops}, 
the EGI Helpdesk \cite{ggus} or Grid configuration repository (GOCDB) \cite{gocdb} and the 
coordination of the of  the external provision of software required by EGI to 
form the production infrastructure. For the later it maintained a production--ready 
Grid computing middleware distribution called {\sl UMD} (Unified Middleware Distribution) and 
defined general and component specific quality criteria to be applied to software 
components made available through this repository.
The role of {\sl UMD} was to enforce the fulfilment of this set of quality
criteria \cite{egi-qc} in the software being delivered (i.e.,
{\sl EMI} and {\sl Globus}) \cite{mario}. {\sl UMD} is still being used and
deployed in the European scientific e--Infrastructures under a follow--up
project, the EGI--Engage (Engaging the EGI Community towards an Open Science
Commons) \cite{cordis:egi-engage}. This distribution is currently complemented
by a Cloud--specific one called {\sl CMD} (Cloud Middleware Distribution).

@@

>>>>

The DataGrid \cite{cordis:datagrid} project (Jan 2001 -- Dec 2003) 
had as main goal to develop the software to provide basic grid functionality
and associated management tools for a large scale testbed for demonstration projects in three
specific areas of science, such as Particle physics, Bioinformatics and Earth observation.
%was devoted to develop the first European infrastructure for Grid computing. 
The project brought together 21 academic and industry partners, from 15 different
countries. 
%There were 3 application areas for requirement gathering; Particle physics, Bioinformatics and Earth observation. 
The project delivered its own software distribution, named EDG (EU DataGrid), strongly 
based on Globus Middleware Services \cite{globus}. The project was organized in 12 workpages, from 
which 5 were devoted to software development coordinated by an 
Architecture Task Force, supervising the overall design and technical consistency 
of the developments.

@@

The source code was available at a private, centralized version control system.
The code was not validated by any Quality Assurance policy, but passed through a manual
certification procedure at the time of the release. This procedure tried to
improve the reliability of the software components by applying a set of
requirements [ref] to be fulfilled at each stage of the  development lifecycle (integration, certification,
pre--production, production), carried out by different teams. However, with 
EGEE--II, the developers started to use the ETICS system \cite{etics}, solution aiming to develop
an automatic build system for Grid middleware.

@@

EGI--InSPIRE not only ensured the coordination of the operation of a secure, reliable 
European-wide production grid infrastructure federated from national grid initiatives 
that is integrated and interoperates with other grids worldwide but also foreseed the 
maintenance and development of operational tools, like the Operations Portal \cite{egi-ops}, 
the EGI Helpdesk \cite{ggus} or Grid configuration repository (GOCDB) \cite{gocdb} and the 
coordination of the external provision of software required by EGI to 
form the production infrastructure. For the later it maintained a production--ready 
Grid computing middleware distribution called {\sl UMD} (Unified Middleware Distribution) and 
defined general and component specific quality criteria to be applied to software 
components made available through this repository.
The role of {\sl UMD} was to enforce the fulfilment of this set of quality
criteria \cite{egi-qc} in the software being delivered (i.e.,
{\sl EMI} and {\sl Globus}) \cite{mario}. {\sl UMD} is still being used and
deployed in the European scientific e--Infrastructures under a follow--up
project, the EGI--Engage (Engaging the EGI Community towards an Open Science
Commons) \cite{cordis:egi-engage}. This distribution is currently complemented
by a Cloud--specific one called {\sl CMD} (Cloud Middleware Distribution).

@@

\subsection{Features on software reliability in EC-projects}

In the following we summarize the features on software reliability in the aforementioned EC-projects.
Table \ref{tab:feat} shows the various projects 
with their features: the symbol $\surd$ crosses project and feature.

\begin{table*}[!h]
%% increase table row spacing, adjust to taste
\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
\caption{Features on software reliability in EC-projects}
\label{tab:feat}
\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
\begin{tabular}{llllllll}
\hline
\hline
Feature & DataGrid & EGEE I, II, III & ETICS 1, 2 & EMI & EGI-Inspire & EGI Engage & INDIGO-DataCloud\\
\hline
\hline
Architecture Task Force&$\surd$&$\surd$&$\surd$&$\surd$&$\surd$&$\surd$&$\surd$\\
Communication Handling&$\surd$&$\surd$&$\surd$&$\surd$&$\surd$&$\surd$&$\surd$\\
Requirements Handling&$\surd$&$\surd$&$\surd$&$\surd$&$\surd$&$\surd$&$\surd$\\
Agile Methodologies&$\surd$&$\surd$&$\surd$&$\surd$&$\surd$&$\surd$&$\surd$\\
Buu/Test Management Procedure&$\surd$&$\surd$&$\surd$&$\surd$&$\surd$&$\surd$&$\surd$\\
Manual Certification Procedure&$\surd$&$\surd$&&$\surd$&&&\\
Automatic Certification Tool&&&$\surd$&&$\surd$&$\surd$&$\surd$\\
Auto-generated Documentation&&$\surd$&$\surd$&$\surd$&$\surd$&$\surd$&$\surd$\\
Software Product Metrics&$\surd$&$\surd$&$\surd$&&$\surd$&$\surd$&$\surd$\\
Quality Criteria Definition&&&$\surd$&$\surd$&$\surd$&$\surd$&$\surd$\\
\hline
\hline
\end{tabular}
\end{table*}
